---
title: "How We Feel Analysis: How is Behavior Affected by Testing?"
author: "Andy Shi"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

Our goal is to see how receiving a SARS-CoV-2 test changes one's
behavior. We will focus on the time course of the behavior two weeks
before and two weeks after receiving a test and compare how this differs
for those tested as SARS-CoV-2(+) or SARS-CoV-2(-) and those predicted
to be SARS-CoV-2(+) or SARS-CoV-2(-).

# Setup and Data Munging

```{r libraries, message = FALSE, warning = FALSE}
library(dplyr)
library(ggplot2)
library(broom)
library(knitr)
library(purrr)
library(data.table)
library(tidyr)
```

First, we define helper functions to recode the data and to collapse the
data to a per-user summary.

```{r recode}
recode_df_helper <- function(df, varname, true_val, false_val) {
  varname_new <- paste0(varname, "_NEW")
  df[get(varname) == true_val, (varname_new) := 1L]
  df[get(varname) == false_val, (varname_new) := 0L]
}

default_tf <- c("True", "False")
recode_lst <- list(
  cancelled_medical_appointment = c("yes", "no"),
  combined_stayed_home = default_tf,
  combined_left_for_work = default_tf,
  combined_left_for_other = default_tf,
  protective_measures_face_mask = default_tf,
  protective_measures_other_face_covering = default_tf,
  protective_measures_social_distancing = default_tf)

recode_df <- function(df, recode_lst) {
  for (i in seq_along(recode_lst)) {
    behav_var <- names(recode_lst)[i]
    true_val <- recode_lst[[behav_var]][1]
    false_val <- recode_lst[[behav_var]][2]
    recode_df_helper(df, behav_var, true_val, false_val)
  }
  df[, `:=` (
      mask_wearing = pmax(protective_measures_face_mask_NEW,
                          protective_measures_other_face_covering_NEW,
                          na.rm = TRUE),
      days_since_test_numeric = as.double(days_since_test,
                                          units = "days"))]
}

```

We then collapse across session_ids by days since test, tested/predicted
status, and positive/negative status. For most of the variables, we take
the mean: the proportion of check-ins on that day where a behavior was
practiced. For the estimated number of people in contact, we take the
median.

```{r timecourse_collapse}
timecourse_collapse <- function(df) {
  timecourse1 <- df[,
      lapply(.SD, mean, na.rm = TRUE),
      by = .(days_since_test, tested_predicted_indicator, positive),
      .SDcols = c("cancelled_medical_appointment_NEW",
                  "combined_stayed_home_NEW",
                  "combined_left_for_work_NEW",
                  "combined_left_for_other_NEW",
                  "protective_measures_social_distancing_NEW",
                  "mask_wearing")]

  timecourse2 <- df[,
    .(nobs = .N,
      estimate_people_contact =
        median(estimate_people_contact, na.rm = TRUE)),
    by = .(days_since_test, tested_predicted_indicator, positive)]

  timecourse_final <-
    merge(timecourse1, timecourse2,
          by = c("days_since_test", "tested_predicted_indicator",
                "positive"))
  return(timecourse_final)
}
```

Below, we munge the tested vs. untested individuals.


```{r munge_test_untested}
df_tested <- readRDS("df_tested.rds")
df_untested <- readRDS("df_untested.rds")

df_test_untest <- rbind(df_tested, df_untested, fill = TRUE)
rm(df_untested)
recode_df(df_test_untest, recode_lst)
timecourse_test_untest <- timecourse_collapse(df_test_untest)
rm(df_test_untest)
```

Below, we munge the tested vs. predicted individuals. Some individuals
who were tested were also predicted, so we remove those from the
analysis.

```{r munge_test_pred}
df_predicted <- readRDS("df_predicted.rds")
df_tested_session_ids <- df_tested[, .(session_id)]
# remove people who were tested
df_predicted_filtered <- df_predicted[!df_tested, ]
rm(df_predicted)

df_test_pred <- rbind(df_tested, df_predicted_filtered, fill = TRUE)
rm(df_predicted_filtered, df_tested)
recode_df(df_test_pred, recode_lst)
timecourse_test_pred <- timecourse_collapse(df_test_pred)
rm(df_test_pred)
```

We then do some more recoding of the variables and also make a long
version of the data.

```{r timecourse_recode}
timecourse_test_untest[is.na(positive), positive := 3]
timecourse_test_untest[, `:=`
  (positive_fct = factor(positive, levels = c(0, 1, 3),
                         labels = c(" Negative", " Positive", "")),
   tested_pred_fct = factor(tested_predicted_indicator))]
timecourse_test_untest[,
  test_pred_res := interaction(tested_pred_fct, positive_fct,
                               sep = "", lex.order = TRUE, drop = TRUE)]

timecourse_test_pred[, `:=`
  (positive_fct = factor(positive, levels = c(0, 1),
                         labels = c(" Negative", " Positive")),
   tested_pred_fct = factor(tested_predicted_indicator))]
timecourse_test_pred[,
  test_pred_res := interaction(tested_pred_fct, positive_fct,
                               sep = "", lex.order = TRUE, drop = TRUE)]

vars_touse <- c("cancelled_medical_appointment_NEW",
                "combined_stayed_home_NEW",
                "combined_left_for_work_NEW",
                "combined_left_for_other_NEW",
                "protective_measures_social_distancing_NEW",
                "mask_wearing",
                "estimate_people_contact")
vars_touse_pretty <- c("Canceled Medical Appt",
                       "Stayed Home",
                       "Left for Work",
                       "Left for Other",
                       "Social Distanced",
                       "Wore Mask",
                       "estimate_people_contact")
timecourse_long_test_untested <- timecourse_test_untest %>%
  select(-positive, -positive_fct, -tested_pred_fct,
         -tested_predicted_indicator) %>%
  pivot_longer(-c(days_since_test, nobs,
                  test_pred_res),
               names_to = "behavior_var",
               values_to = "mean_users") %>%
  mutate(behavior_fct =
         factor(behavior_var, levels = vars_touse,
                labels = vars_touse_pretty))

timecourse_long_test_pred <- timecourse_test_pred %>%
  select(-positive, -positive_fct, -tested_pred_fct,
         -tested_predicted_indicator) %>%
  pivot_longer(-c(days_since_test, nobs,
                  test_pred_res),
               names_to = "behavior_var",
               values_to = "mean_users") %>%
  mutate(behavior_fct =
         factor(behavior_var, levels = vars_touse,
                labels = vars_touse_pretty))
```

# Plots

We first define the colors, line type, and shapes we want to use, and
then plot the behavior time course for the variables we considered.

```{r styles}
neg_color <- "#79b5d9"
pos_color <- "#f16347"
color_vals1 <- c("Untested" = "#000000",
                 "Tested Negative" = neg_color,
                 "Tested Positive" = pos_color)
lty_vals1 <- c("Untested" = "dashed",
               "Tested Negative" = "solid",
               "Tested Positive" = "solid")
shape_vals1 <- c("Untested" = 17,
                 "Tested Negative" = 20,
                 "Tested Positive" = 20)

color_vals2 <- c("Predicted Negative" = neg_color,
                 "Predicted Positive" = pos_color,
                 "Tested Negative" = neg_color,
                 "Tested Positive" = pos_color)
lty_vals2 <- c("Predicted Negative" = "dashed",
               "Predicted Positive" = "dashed",
               "Tested Negative" = "solid",
               "Tested Positive" = "solid")
shape_vals2  <- c("Predicted Negative" = 17,
                 "Predicted Positive" = 17,
                 "Tested Negative" = 20,
                 "Tested Positive" = 20)
```

## Behavior Timecourse Plots

```{r behavior_timecourse_plt}

behavior_timecourse_test_untested <- timecourse_long_test_untested %>%
  filter(behavior_var != "estimate_people_contact") %>%
  ggplot(aes(x = days_since_test, y = mean_users,
             color = test_pred_res, linetype = test_pred_res)) +
  geom_line() +
  xlab("Days since test") +
  ylab("Prop. of Users Engaging") +
  facet_wrap(vars(behavior_fct)) +
  scale_color_manual(name = "Key", values = color_vals1) +
  scale_linetype_manual(name = "Key", values = lty_vals1) +
  geom_vline(xintercept = 0, lty = 2, color = "black", lwd = 1) +
  theme_bw() +
  theme(text = element_text(size = 15),
        legend.position = "bottom",
        panel.spacing = unit(2, "lines")) +
  ggtitle("Change in Behavior After Testing")
ggsave("figures/behavior_timecourse_test_untested.pdf",
       behavior_timecourse_test_untested, width = 9, height = 6)

behavior_timecourse_test_pred <- timecourse_long_test_pred %>%
  filter(behavior_var != "estimate_people_contact") %>%
  ggplot(aes(x = days_since_test, y = mean_users,
             color = test_pred_res, linetype = test_pred_res)) +
  geom_line() +
  xlab("Days since test") +
  ylab("Prop. of Users Engaging") +
  facet_wrap(vars(behavior_fct)) +
  scale_color_manual(name = "Key", values = color_vals2) +
  scale_linetype_manual(name = "Key", values = lty_vals2) +
  geom_vline(xintercept = 0, lty = 2, color = "black", lwd = 1) +
  theme_bw() +
  theme(text = element_text(size = 15),
        legend.position = "bottom",
        panel.spacing = unit(2, "lines")) +
  ggtitle("Change in Behavior After Testing")
ggsave("figures/behavior_timecourse_test_pred.pdf",
       behavior_timecourse_test_pred, width = 9, height = 6)
```

## Estimated Number of Contacts

For the estimated number of contacts, we plot the data on a log10 scale
(technically log10(1 + x)). We fit a loess curve to the points to smooth
out the trend.

```{r est_contact_plt}
est_contact_logscale_test_untest <- timecourse_test_untest %>%
  ggplot(aes(x = days_since_test,
             y = log10(1 + estimate_people_contact),
             color = test_pred_res, linetype = test_pred_res,
             shape = test_pred_res)) +
  geom_jitter(alpha = 0.5, width = 0.2, height = 0) +
  geom_smooth(method = "loess") +
  geom_vline(xintercept = 0, lty = 2, color = "black", lwd = 2) +
  xlab("Days since test") + ylab("log10(1 + Med. # of people contacted)") +
  scale_color_manual(name = "Key", values = color_vals1) +
  scale_linetype_manual(name = "Key", values = lty_vals1) +
  scale_shape_manual(name = "Key", values = shape_vals1) +
  theme_bw() +
  theme(text = element_text(size = 15)) +
  ggtitle("Median Number of People Contacted Before/After Test")
ggsave("figures/est_contact_logscale_test_untest.pdf",
       est_contact_logscale_test_untest, width = 8, height = 6)

est_contact_logscale_test_pred <- timecourse_test_pred %>%
  ggplot(aes(x = days_since_test,
             y = log10(1 + estimate_people_contact),
             color = test_pred_res, linetype = test_pred_res,
             shape = test_pred_res)) +
  geom_jitter(alpha = 0.5, width = 0.2, height = 0) +
  geom_smooth(method = "loess") +
  geom_vline(xintercept = 0, lty = 2, color = "black", lwd = 2) +
  xlab("Days since test") + ylab("log10(1 + Med. # of people contacted)") +
  scale_color_manual(name = "Key", values = color_vals2) +
  scale_linetype_manual(name = "Key", values = lty_vals2) +
  scale_shape_manual(name = "Key", values = shape_vals2) +
  theme_bw() +
  theme(text = element_text(size = 15)) +
  ggtitle("Median Number of People Contacted Before/After Test")
ggsave("figures/est_contact_logscale_test_pred.pdf",
       est_contact_logscale_test_pred, width = 8, height = 6)
```

## Appendix: Number of observations

Finally, we show the number of observations in each category during the
window of time considered.

```{r nobs_plt}
nobs_test_untest <- timecourse_test_untest %>%
  ggplot(aes(x = days_since_test, y = nobs,
             color = test_pred_res, linetype = test_pred_res)) +
  geom_line() +
  scale_color_manual(name = "Key", values = color_vals1) +
  scale_linetype_manual(name = "Key", values = lty_vals1) +
  xlab("Days since test") + ylab("Number of responses") +
  scale_y_log10() +
  geom_vline(xintercept = 0, lty = 2, color = "black", lwd = 1) +
  theme_bw() +
  theme(text = element_text(size = 15)) +
  ggtitle("Number of responses over time")
ggsave("figures/nobs_test_untest.pdf", nobs_test_untest,
       width = 8, height = 6)

nobs_test_pred <- timecourse_test_pred %>%
  ggplot(aes(x = days_since_test, y = nobs,
             color = test_pred_res, linetype = test_pred_res)) +
  geom_line() +
  scale_color_manual(name = "Key", values = color_vals2) +
  scale_linetype_manual(name = "Key", values = lty_vals2) +
  xlab("Days since test") + ylab("Number of responses") +
  scale_y_log10() +
  geom_vline(xintercept = 0, lty = 2, color = "black", lwd = 1) +
  theme_bw() +
  theme(text = element_text(size = 15)) +
  ggtitle("Number of responses over time")
ggsave("figures/nobs_test_pred.pdf", nobs_test_pred,
       width = 8, height = 6)
```


# Session Info

```{r}
sessionInfo()
```
